---
title: "More data Wrangling with joins and tidyR"
#author: "Insert Name"
date: "Math 241, Week 4"
output:
  pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
# Do not modify this chunk.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r libraries, message = FALSE}
# it's good practice to check that all the packages required are loaded and installed
libs <- c('tidyverse','knitr','viridis','mosaicData','babynames','mdsr','Lahman','nycflights13')
for(l in libs){
  if(!require(l,character.only = TRUE, quietly = TRUE)){
    message( sprintf('Did not have the required package << %s >> installed. Downloading now ... ',l))
    install.packages(l) 
  }
  library(l, character.only = TRUE, quietly = TRUE)
}
```

## Goals of this in-class activity:

- Practice data wrangling and joins with tidyR

## Notes:

* Be prepared to ask for help from me, Tory, and your classmates!

## Problem 1 (Medium): 
Consider the number of home runs hit (HR) and home runs allowed (HRA) for the Chicago Cubs (CHN) baseball team. Reshape the Teams data from the `Lahman` package into “long” format and plot a time series conditioned on whether the HRs that involved the Cubs were hit by them or allowed by them.

```{r}

Teams %>% 
  filter(teamID == "CHN") %>%
  select(yearID, HR, HRA) %>%
  pivot_longer(-yearID, names_to = "type", values_to = "home_runs") %>%
  ggplot(aes(x = yearID, y = home_runs, color = type)) +
  geom_point() +
  geom_line()
```

## Problem 2 (Medium): 
Use the `nycflights13` package and the `flights` and `planes` tables to answer the following questions:

a. How many planes have a missing date of manufacture?

```{r}
library(nycflights13)
planes2 <- select(planes, tailnum, year, manufacturer)
flights2 <- select(flights, tailnum)
nyc_flights <- left_join(planes2, flights2)
nyc_flights2 <- nyc_flights %>%
  filter(is.na(year)) %>%
  distinct(tailnum)
nrow(nyc_flights2)
```

There are 70 airplanes with a missing date of manufacture.


b. What are the five most common manufacturers?

```{r}
nyc_flights %>%
  select(manufacturer, tailnum, year) %>%
  unique() %>%
  group_by(manufacturer) %>%
  summarize(count = n()) %>%
  arrange(desc(count))
```


## Problem 3 (Medium): 
Use the `nycflights13` package and the `flights` and `planes` tables to answer the following questions:

a. What is the oldest plane (specified by the `tailnum` variable) that flew from
New York City airports in 2013?

```{r}
planes2 <- dplyr::select(planes, tailnum, year)
flights2 <- dplyr::select(flights, tailnum)
nyc_flights <- left_join(planes2, flights2) %>%
  arrange(year) %>%
  unique()
head(nyc_flights)
```

N381AA, manufactured in 1956, is the oldest plane that flew from NYC in 2013.

b. How many airplanes that flew from New York City are included in the
planes table?

```{r}
nyc_flights2 <- distinct(nyc_flights)
nrow(nyc_flights2)
```

There are 3322 unique airplanes.

## Problem 4 (Medium): 
The `knitr` package allows the analyst to display nicely formatted tables and results when outputting to pdf files. Use the following code chunk as an example to create a similar display for the `penguins` dataset, in the `palmerpenguins` package, instead (you can model penguins' `body_mass_g` as a function of their `flipper_length_mm` and `sex`):

```{r}
mod <- broom::tidy(lm(cesd ~ mcs + sex, data = HELPrct))
knitr::kable(
  mod,
  digits = c(0, 2, 2, 2, 4),
  caption = "Regression model from HELP clinical trial.",
  longtable = TRUE
)
```

```{r}
library(palmerpenguins)
mod <- broom::tidy(lm(body_mass_g ~ flipper_length_mm + sex, data = penguins))
knitr::kable(mod, digits = c(0, 1, 1, 1, 4), longtable = TRUE)
```
