---
title: "Practice ggplot2"
#author: "Insert Name"
date: "Math 241, Week 1"
output:
  pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
# Do not modify this chunk.
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
```

```{r libraries, message = FALSE}
# it's good practice to check that all the packages required are loaded and installed
libs <- c('tidyverse','dplyr','ggplot2','knitr','viridis','mdsr', 'macleish','babynames')
for(l in libs){
  if(!require(l,character.only = TRUE, quietly = TRUE)){
    message( sprintf('Did not have the required package << %s >> installed. Downloading now ... ',l))
    install.packages(l) 
  }
  library(l, character.only = TRUE, quietly = TRUE)
}
```

## Goals of this in-class activity:

- Practice creating and refining graphs with `ggplot2`.
- Consider the strengths and weaknesses of various `geom`s and `aes`thetics for telling a data story.

## Notes:

* When creating your graphs, consider context (i.e. axis labels, title, ...)!
* If I provide partially completed code, I will put `eval = FALSE` in the chunk.  Make sure to change that to `eval = TRUE` once you have completed the code in the chunk.
* Be prepared to ask for help from me, Tory, and your classmates!  We scratch the surface of `ggplot2` in class.  But I encourage you to really dig in and make your graphs your own (i.e. don't rely on defaults).

## Problem 1 (Easy): 

Consider the following data graphic.
```{r}
ggplot(mtcars, aes(x = disp, y = mpg, color = factor(am))) +
geom_point() +
  labs(x = "Displacement (cu. in.)",
       y = "Miles per gallon",
       color = "Manual Transmission",
       title = "Fuel efficiency vs. engine displacement") +
  theme_classic()
```

The am variable takes the value 0 if the car has automatic transmission and 1 if
the car has manual transmission. How could you differentiate the cars in the
graphic based on their transmission type?


## Problem 2 (Easy): 

Angelica Schuyler Church (1756–1814) was the daughter of New York Governer Philip Schuyler and sister of Elizabeth Schuyler Hamilton. Angelica, New York was named after her. 

Using the `babynames` package generate a plot of the reported proportion of babies born with the name Angelica over time and interpret the figure.

```{r}
data(babynames) # this will explicitly ask R to load the babynames dataset to your environment
angelica <- filter(babynames, name == "Angelica")
ggplot(data = angelica, aes(x = year, y = prop, color = sex)) +
geom_line() +
  labs(x = 'Year', y = '% Angelica', color = 'Biological Sex') +
  theme_classic()
```
We see a huge increase in the proportion of girls named Angelica.

## Problem 3 (Medium): 

The `macleish` package contains weather data collected every 10 minutes in 2015 from two weather stations in Whately, MA.


```{r}
whately_2015 <- whately_2015 %>% mutate(Date = as.Date(when))

ggplot(data = whately_2015, aes(x = Date, y = temperature)) +
  geom_line(size = 0.3, color = "darkgray") +
  labs(y = "Temperature (degrees C)", title = "Temperature measurements at Macleish Field Station") +
  geom_smooth() +
  scale_x_date() +
  theme_classic()
```

Using ggplot2, create a data graphic that displays the average temperature over
each 10-minute interval (temperature) as a function of time (when).

## Problem 4 (Medium): 

The data set `MLB_teams` in the `mdsr` package contains information about Major League Baseball teams from 2008–2014. There are several quantitative and a few categorical variables present. See how many variables you can illustrate on a single plot in R. The current record is 7.
(Note: **This is not good graphical practice**—it is merely an exercise to help you
understand how to use visual cues and aesthetics!)


```{r, fig.height=12, fig.width = 10}

ggplot(MLB_teams) +
  geom_point(aes( x = payroll / 1000000, y = WPct, 
                  color = teamID, size = metroPop, shape = lgID, alpha = normAttend)) +
  facet_wrap(~yearID, ncol = 1) +
  xlab("Payroll (Millions of USD)") +
  ylab("Team Winning Percentage") +
  theme_classic()
```


## Problem 5 (Medium): 
Use the `MLB_teams` data in the `mdsr` package again to create an informative data graphic that illustrates the relationship between winning percentage and payroll in context.

```{r}
options(scipen = 999) # this will suppress scientific notation
ggplot(data = MLB_teams, aes(
  x = payroll, y = WPct,
  alpha = yearID,
  color = lgID, size = metroPop)) +
  geom_point() +
  geom_smooth() +
  xlab("Payroll (Millions of USD)") +
  ylab("Team Winning Percentage") +
  theme_classic()

```

## Problem 6 (Hard): 
Use the function `make_babynames_dist()` in the `mdsr` package to recreate the “Deadest Names” graphic from [FiveThirtyEight](https://fivethirtyeight.com/features/how-to-tell-someones-age-when-all-you-know-is-her-name/).

```{r}

babynames_dist <- make_babynames_dist()
glimpse(babynames_dist)

babynames_dist %>%
  filter(year >= 1900) %>%
  group_by(name, sex) %>%
  summarize(
  N = n(),
  total_est_alive_today = sum(est_alive_today),
  total = sum(n)) %>%
mutate(pct_dead = 1 - (total_est_alive_today / total)) %>%
  filter(total > 50000) %>%
  arrange(desc(pct_dead)) %>%
  head(20) %>%
  ggplot(aes(x = reorder(name, pct_dead), y = pct_dead, fill = sex)) +
  geom_bar(stat = "identity") +
  geom_text(
  aes(
  y = pct_dead + 0.05,
  label = paste(round(pct_dead * 100, 1), "%")
  )
  ) +
  coord_flip() +
  ggtitle("Deadest Names",
  subtitle =
  "Estimated % of Americans with a given name\nborn since 1900 who were dead as of Jan. 1, 2017"
  ) +
  scale_x_discrete(NULL) +
  scale_y_continuous(NULL) +
  scale_fill_manual(values = c("#f6b900", "#008fd5")) +
  theme_classic()
```

