---
title: "gganimate - activity"
editor: visual
editor_options: 
  chunk_output_type: console
---

# Introduction

Visualizing how civil liberties have evolved over time.

```{r}
#| label: load-packages
#| echo: false
#| message: false
#| warning: false
library(countdown)
library(tidyverse)
library(gt)
library(readxl)
library(gganimate)
library(gifski)
library(knitr)
library(kableExtra)
library(palmerpenguins)
library(transformr)
library(datasauRus)

```

## Data

Freedom House has analyzed challenges to freedom and advocated for greater political rights and civil liberties in a variety of publications over the years. They produce an ordinal variable from 1 to 7- from more to less civil liberties (More about the data)\[https://freedomhouse.org/reports/publication-archives\].

```{r}
#| label: get-data
#| cache: true
#| echo: false
#| message: false
#| warning: false
#| out-width: "80%"

freedom <- read_csv(here::here("freedom.csv"), na = "-")

freedom_ranked <- freedom %>%
  select(country, contains("cl_")) %>%
  pivot_longer(
    cols = -country,
    names_to = "year",
    values_to = "civil_liberty",
    names_prefix = "cl_",
    names_transform = list(year = as.numeric)
  ) %>%
  group_by(year) %>%
  mutate(rank_in_year = rank(civil_liberty, ties.method = "first")) %>%
  ungroup()

```

# Your first interactive visualization!

-   **Demo:**: create a simple static line plots of your country of choice civl liberty score over time

```{r}
#| label: simple-line-plot
#| eval: false

freedom_ranked %>%
  filter() %>% 
  ggplot()

```

-   **Demo:** Build on your previous plot to create an interactive version of the figure. Hint: use `transition_reveal()`

# Visualizing multiple countries at once!

In this part, rather than focusing on one country's civil liberty score over time, we want to look at countries' ranking.

We have a long list of countries, and we may want to focus on visualizing the ranking of a limited number of countries over time.

```{r}
#| label: subset-countries
#| eval: false
countries_to_plot <- freedom %>%
  rowwise() %>%
  mutate(sd = sd(c_across(contains("cl_")), na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(sd)) %>%
  relocate(country, sd) %>%
  slice_head(n = 15) %>%
  pull(country)

freedom_to_plot <- freedom %>%
  filter(country %in% countries_to_plot) %>%
  drop_na()

freedom_ranked <- freedom %>%
  select(country, contains("cl_")) %>%
  pivot_longer(
    cols = -country,
    names_to = "year",
    values_to = "civil_liberty",
    names_prefix = "cl_",
    names_transform = list(year = as.numeric)
  ) %>%
  group_by(year) %>%
  mutate(rank_in_year = rank(civil_liberty, ties.method = "first")) %>%
  ungroup()

```

-   **Demo:** Create a simple static bar plot showing a country's `civil_liberty` as a function of a country's `rank` for the years 1990, 2000, 2010.

```{r}
#| label: geom-col
#| eval: false

# add code here

# hint: to label each bar with country name, you can use 

geom_text(
  hjust = "right", 
  aes(label = country), 
  x = -1)

```

## Interactive figure

-   **Demo:** Build on your previous plot to create an interactive version of the figure. Specifically, you want to see the ranking over the whole time series. Hint: use `transition_time()`

```{r}

#| label: interactive-geom-col
#| eval: false

# hint: make sure all the years are included this time


# add code here

# hint: label each year with   

geom_text(
    x = 5, y = 1,
    hjust = "left",
    aes(label = as.character(year)),
    size = 10
  )

```
