---
title: "Practicing Regular Expressions"
subtitle: "Math 241, Week 8"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
#Load tidyverse
library(tidyverse)
library(mdsr)
```


## Problem 1 (Easy): 

Use the `Macbeth_raw` data from the `mdsr` package to answer the following questions:

a. Find all of the adjectives in Macbeth that end in more or less


```{r}
data(Macbeth_raw)
lines <- strsplit(Macbeth_raw, "\r\n")[[1]]
words <- strsplit(paste(lines, collapse = " "), " ")[[1]]
grep("[a-z]+(more|less)$", words, value = TRUE)
```


b. Find all of the lines containing the stage direction Exit or Exeunt.

```{r}
lines <- strsplit(Macbeth_raw, "\r\n")[[1]]
grep("(Exit|Exeunt)", lines, value = TRUE)
```

## Problem 2 (Medium): 

a. Use regular expressions to determine the number of speaking lines in the Complete Works of William Shakespeare. Here, we care only about how many times a character speaks—not what they say or for how long they
speak.

```{r}
shakespeare_url <- "https://www.gutenberg.org/files/100/100-0.txt"
shakespeare <- RCurl::getURL(shakespeare_url) %>% str_split("\r\n") %>%
  pluck(1)
length(shakespeare)
```

```{r}
speaker_lines_pattern <- "[A-Z ]+\\." # I want any number of uppercase characters followed by a period
speaker_lines <- grep(speaker_lines_pattern,shakespeare)
length(speaker_lines)
```

b. Make a bar chart displaying the top 100 characters with the greatest number of lines. Hint you may want to use either the `stringr::str_extract` or `strsplit` function here.

```{r}
speakers <- str_extract(shakespeare, speaker_lines_pattern)
speaker_df <- speakers %>% 
  data.frame(speaker = .) %>% 
  filter(!is.na(speaker)) %>% 
  group_by(speaker) %>% 
  summarize(N = n()) %>%
  arrange(desc(N)) %>% head(100)

ggplot(data = speaker_df, aes(x = reorder(speaker, N), y = N)) +
  geom_bar(stat = "identity") +
  coord_flip()
```


