---
title: "Lab 3"
#author: "Insert Name"
date: "Math 241, Week 3"
output:
  pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
# Do not modify this chunk.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


```

```{r}
libs <- c('tidyverse','knitr','viridis', 'mosaic','mosaicData','babynames', 'Lahman','nycflights13','rnoaa')
for(l in libs){
  if(!require(l,character.only = TRUE, quietly = TRUE)){
    message( sprintf('Did not have the required package << %s >> installed. Downloading now ... ',l))
    install.packages(l) 
  }
  library(l, character.only = TRUE, quietly = TRUE)
}

```



## Due: Friday, February 16th at 8:30am

## Goals of this lab

1. Practice creating functions.
1. Practice refactoring your code to make it better!  Therefore for each problem, make sure to test your functions.


### Problem 1: Subset that R Object

Here are the R objects we will use in this problem (`dats`, `pdxTreesSmall` and `ht`).    

```{r}
library(pdxTrees)
library(mosaicData)

pdxTrees <- get_pdxTrees_parks()
# Creating the objects
dats <- list(pdxTrees  = head(pdxTrees),
             Births2015 = head(Births2015),
             HELPrct = head(HELPrct), 
             sets = c("pdxTrees", "Births2015", 
                      "HELPrct"))

pdxTreesSmall  <- head(pdxTrees)

ht <- head(pdxTrees$Tree_Height, n = 15)

```

a. What are the classes of `dats`, `pdxTreesSmall` and `ht`?

b. Find the 10th, 11th, and 12th values of `ht`.

c. Provide the `Species` column of `pdxTrees` as a data frame with one column.

d. Provide the `Species` column of `pdxTrees` as a character vector.

e. Provide code that gives us the second entry in `sets` from `dats`.

f. Subset `pdxTreesSmall` to only `Douglas-fir` and then provide the `DBH` and `Condition` of the 4th `Douglas-fir` in the dataset.  (Feel free to mix in some `tidyverse` code if you would like to.)




### Problem 2: Function Creation

Figure out what the following code does and then turn it into a function. For your new function, do the following:

* Test it.
* Provide default values (when appropriate).
* Use clear names for the function and arguments.
* Make sure to appropriately handle missingness.
* Generalize it by allowing the user to specify a confidence level.
* Check the inputs and stop the function if the user provides inappropriate values.


```{r}
library(pdxTrees)
thing1 <- length(pdxTrees$DBH)
thing2 <- mean(pdxTrees$DBH)
thing3 <- sd(pdxTrees$DBH)/sqrt(thing1)
thing4 <- qt(p = .975, df = thing1 - 1)
thing5 <- thing2 - thing4*thing3
thing6 <- thing2 + thing4*thing3
```





### Problem 3: Wrapper Function for your `ggplot`


While we (i.e. Math 241 students) all love the grammar of graphics, not everyone else does. So for this problem, we are going to practice creating wrapper functions for `ggplot2`.

Here's an example of a wrapper for a histogram.  Notice that I can't just list the variable name as an argument.  The issue has to do with how many of the `tidyverse` functions evaluate the arguments. Therefore we have to quote (`enquo()`) and then unquote (`!!`) the arguments.  (If you want to learn more, go [here](https://tidyeval.tidyverse.org/).)

```{r}
# Minimal viable product working code
ggplot(data = pdxTrees, mapping = aes(x = DBH)) +
  geom_histogram()

# Shorthand histogram function
histo <- function(data, x){
   x <- enquo(x)
  ggplot(data = data, mapping = aes(x = !!x)) +
    geom_histogram()
}

# Test it
histo(pdxTrees, DBH)
```

a. Edit `histo()` so that the user can set 

* The number of bins
* The fill color for the bars
* The color outlining the bars

b. Write code to create a basic scatterplot with `ggplot2`. Then write and test a function to create a basic scatterplot.

c. Modify your scatterplot function to allow the user to ...

* Color the points by another variable.
* Set the transparency.

d. Write and test a function for your favorite `ggplot2` graph.


### Problem 4: Functioning `dplyr`


a. Take the following code and turn it into an R function to create a **conditional proportions** table.  Similar to `ggplot2`, you will need to quote and unquote the variable names.  Make sure to test your function!

```{r}
pdxTrees %>%
  count(Native, Condition) %>%
  group_by(Native) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup()
```


b.  Write a function to compute the mean, median, sd, min, max, sample size, and number of missing values of a quantitative variable by the categories of another variable.  Make sure the output is a data frame (or tibble).  Don't forget to test your function.


### Problem 5: another `babynames` exercise

Write a function called grab_name that, when given a **name** ***and a year*** as an argument, returns the rows from the `babynames` data frame in the `babynames` package that match that name for that year (and returns an error if that name and year combination does not match any rows). Run the function once with the arguments **Ezekiel and 1883** and once with **Ezekiel and 1983**.

```{r, eval=FALSE}
#' Make sure to switch eval = FALSE to eval = TRUE before knitting!!

grab_name <- function(myname, myyear) {
  # Code your function here
}

```