---
title: "Lab 8"
date: "Math 241, Week 11"
format: html
editor: visual
---

```{r}
#| label: setup
#| message: false
#| echo: false
library(tidyverse)
library(tidymodels)
library(openintro)
library(mdsr)
library(mosaicData)
library(shiny)
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE)
```

## Due: Friday, April 12th at 8:30am

## Goals of this lab

-   In this lab, you will work with observational data to assess the relationship between an exposure and the incidence of a disease.
-   You will practice bootstrap to calculate a 95%CI.
-   You will create an interactive web application with `shiny`.

## Exercise 1:

The `Whickham` data set in the `mosaicData` package includes data on age, smoking, and mortality from a one-in-six survey of the electoral roll in Whickham, a mixed urban and rural district near Newcastle upon Tyne, in the United Kingdom. The survey was conducted in 1972--1974 to study heart disease and thyroid disease. A follow-up on those in the survey was conducted 20 years later. Describe the association between smoking status and mortality in this study. Be sure to consider the role of age as a possible confounding factor.

## Exercise 2:

a)  Calculate the mean age of birthing people from the `Gestation` data set from the `mosaicData` package.

```{r}
```

b)  Use the bootstrap to generate and interpret a 95% confidence interval for the mean age of birthing people.

```{r}

```

## Exercise 3:

Using data from the `palmerpenguins` package, create a Shiny app that displays measurements from the penguins dataframe. Allow the user to select a species or a gender, and to choose between various attributes on a scatterplot. (Hint: examples of similar apps can be found at the Shiny gallery).

```{r}

```
